---
title: "Incarceration Data Visualization"
author: "Vannary Sou"
date: "11/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This report will analyze data from [Vera Institute's dataset on incarceration trends](https://raw.githubusercontent.com/vera-institute/incarceration-trends/master/incarceration_trends.csv). The incarceration trends dataset includes county and jurisdiction level data from 1970-2018 and prison data from 1983-2016 in the U.S. 

For my analysis, the problem domain that I will be focusing on is the prison admission rate based on race and sex (male or female) overtime in various parts of the U.S. as well as the country as a whole. The measures of incarceration I have chosen to analyze are: 

* Year 
  * year
  
* State 
  * state
  
* Total prison admission rate 
  * total_prison_adm_rate
  
* Prison admission rate by race 
  * aapi_prison_adm_rate
  * black_prison_adm_rate
  * latinx_prison_adm_rate
  * native_prison_adm_rate
  * white_prison_adm_rate
  
* Prison admission rate by sex 
  * female_prison_adm_rate
  * male_prison_adm_rate

I chose to analyze these variables to identify trends based on the admission rate for different groups of people in various parts of the U.S. throughout history. 

## Summary Statistics 

A paragraph of summary information, citing at least 5 values calculated from the data

You will share at least 5 relevant values of interest. These will likely be calculated using your DPLYR skills, answering questions such as: 

What is the average value of my variable across all the counties (in the current year)?
Where is my variable the highest / lowest?
How much has my variable change over the last N years?

Feel free to calculate and report values that you find relevant. Again, remember that the purpose is to think about how these measure of incarceration vary by race. 

```{r}

```

## Time Trend Chart

The first chart that you'll create and include will show the trend over time of your variable/topic. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

Show more than one, but fewer than ~10 lines: your graph should compare the trend of your measure over time. This may mean showing the same measure for different locations, or different racial groups. Think carefully about a meaningful comparison of locations (e.g., the top 10 counties in a state, top 10 states, etc.)
You must have clear x and y axis labels,
The chart needs a clear title 
You need a legend for your different line colors, and a clear legend title

In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

```{r}

```


## Variable Comparison Chart

The second chart that you'll create and include will show how two different (continuous) variables are related to one another. Again, think carefully about what such a comparison means, and want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

You must have clear x and y axis labels,
The chart needs a clear title 
If you choose to add a color encoding (not required), you need a legend for your different color and a clear legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

```{r}

```


## Map

The last chart that you'll create and include will show how a variable is distributed geographically. Again, think carefully about what such a comparison means, and want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

Your map needs a title
Your color scale needs a legend with a clear label
Use a map based coordinate system to set the aspect ratio of your map (see reading)
Use a minimalist theme for the map (see reading)
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

```{r}

```

